Peridynamic Beams: \\ A Non-ordinary, State Based Model

### 2014-01-30 BeamPaper.tex
add subsection on gradient local beams

This paper begins to remedy that lack in the simplest case by presenting a peridynamic equivalent to an Euler-Bernoulli beam.
In addition to directly modeling a thin beam in bending, the simple beam case lays the theoretical framework for more complex peridynamic beam, plate, and shell bending models.
Because many analyses of interest are partly or wholly comprised of these types of features, their development is an important addition to the capabilities of peridynamic analysis.

The second section of this paper provides a brief introduction to peridynamics, including state based models.
The third section presents the state based beam model and demonstrates equivalence to classical Euler-Bernoulli beam theory in the limit of shrinking nonlocality.
The fourth section compares the model to Eringen's nonlocal elasticity for small peridynamic horizons.
The fifth section demonstrates the beam model with simple test cases.
%
\section{Peridynamics}
\label{sec:PDintro}
The term \textit{peridynamic} was coined by Silling in \cite{silling2000reformulation} and alludes to the fact that the force at a point is affected by nearby material configuration.
Weakly nonlocal continuum mechanics models date to the work of Kr\"oner \cite{kroner1967elasticity} and of Eringen and Edelen \cite{eringen1972nonlocal}, and include higher-order displacement derivatives.
These higher-order derivatives impose greater stronger continuity requirements than the classical model. 
By contrast, the peridynamic model is strongly nonlocal and casts material behavior at a point as the \textit{integral} equation 
%
\begin{equation}
\label{eq:PDCoPV}
\rho(\mathbf{x})\ddot{\mathbf{u}}(\mathbf{x}) = \int_\Omega \mathbi{f}(\mathbf{x},\mathbf{q}) dV_\mathbf{q}  + \mathbf{b}(\mathbf{x}) \notag
\end{equation}
%
rather than the classical \textit{differential} equation.
Instead of the divergence of stress, we have the integral of a ``force" function $\mathbi{f}$ of the position vector $\mathbf{x}$ and the position vector $\mathbf{q}$ of a point within the body domain $\Omega$. 
This force function may depend on \(\mathbf{x}\), \(\mathbf{q}\), their deformed positions, the original and deformed positions of other points in \(\Omega\), history, etc.

@@

@@

@@

\subsection{Weighting Function and Perfect Plasticity}
The weighting function \(\omega(\xi)\) describes the relative contribution of each bond-pair, and can be defined according to physical or mathematical considerations. 
Consider an Euler beam in bending with curvature \(\kappa\). 
Fibers running parallel to the neutral axis of the beam are stretched in proportion to their distance from the neutral axis, with strain \(\epsilon = y\kappa\). 
If the fibers are linearly elastic, then the axial stress at each location is \(\sigma = E\epsilon = Ey\kappa\), and the contribution to supported moment \(dM = \kappa E y^2 dA\). 
By comparing the formulations for the moments carried by the Euler beam in \cref{fig:EulerBending} and those of the bond-pair beam in \cref{fig:BPBending}, we see some definite parallels.
%
\begin{figure}[h]
\subinputfrom{\diagrampath}{EulerBending.eps_tex}
\caption{Euler beam moment contribution}
\label{fig:EulerBending}
\end{figure}
%%
\begin{figure}[h]
\subinputfrom{\diagrampath}{BondPairBending.eps_tex}
\caption{Bond-pair moment contribution}
\label{fig:BPBending}
\end{figure}
%
\begin{align}
M_\text{classical}&=\int_{-\frac{t}{2}}^{\frac{t}{2}} \sigma \; y \; dA && = \int_{-\frac{t}{2}}^{\frac{t}{2}} E \kappa \; y^2 \; b(y) dy\notag \\
%
M_\text{PD}&=\int_{-\delta}^{\delta} \vstate{T}{}{\xi}\; \xi \; d\xi & = \int_{-\delta}^{\delta} \alpha \frac{\sin(\Delta\theta)}{|\xi|} \; \xi \; \omega(\xi) d\xi\: & \approx \int_{-\delta}^{\delta} \alpha \kappa |\xi| \; \omega(\xi) d\xi\notag
\end{align}
%
The term \(y\) is the distance from the beam's neutral axis and \(b(y)\) is the width of the beam at that distance from the neutral axis. 
The similarity between classical and peridynamic moment formulations suggests a possible formulation for the weighting function:
%
\begin{equation}
\omega(\xi) = |\xi| b\left(y\right) \text{at}\; y=\frac{\xi}{\delta} \frac{t}{2}\notag
\end{equation}
%
\begin{figure}[h]
\subinputfrom{\diagrampath}{WeightProfile.eps_tex}
\caption{A cross-section dependent weight function}
\label{fig:WeightProfile}
\end{figure}
%
This weight function analogizes the relative contributions of bond pairs of different lengths to the relative contributions of fibers at different distances from the centerline. 
An example is shown in \cref{fig:WeightProfile}.
While this weighting function offers no advantages over a uniform weight function in the case of the linearly elastic beam, it offers a way to model advancing plasticity.

@@

@@

\section{Numerical Simulation}
\subsection{Discretized Model}
For analysis purposes, it is often useful to represent a continuum part with a discrete version. 
Discretizing the bond-pair model is primarily matter of exchanging integrals for sums. 
%
\begin{align}
%\label{eq:discreteBeamw}
w = \alpha \left[1+\cos(\theta(\vstate{Y}{}{\xi},\vstate{Y}{}{\mathbf{-\xi}})) \right] \approx \frac{\alpha}{2}\left(\frac{y(x+\xi)-2y(x)+y(x-\xi)}{\xi}\right)^2 \notag
\end{align}
%
\begin{align}
\label{eq:discretebeam}
\alpha &= \frac{c\; \Delta x}{m} ;\; c= EI ;\; m=\sum_{i=1}^n \omega(\xi_i)\xi_i^2 \implies \nonumber \\
W&=\Delta x \sum_{i=1}^n \frac{EI}{2}\left(\frac{y(x+\xi_i)-2y(x)+y(x-\xi_i)}{\xi_i}\right)^2\notag
\end{align}
%

Discretizing the bond-pair model as proposed above requires that nodes be evenly spaced along the entire beam, otherwise the displacement \(y(x-\xi_i)\) is ill-defined. 
For this reason, the discretization does not allow for areas of higher and lower ``resolution''. 
\subsection{Results}
The simplest test case for this model is a linear-elastic beam with a square profile.
The models below were constrained to purely vertical movement of each point in order to reduce the total number of degrees of freedom, speeding evaluation and improving convergence.
For comparison, equivalent models are created and analyzed in Abaqus 6.12 to verify simple cases.
Even a coarse discretization successfully reproduces the elastically deformed beam shape in \cref{fig:eppt_g100}.

@@

\section{Conclusion}

>>>>

This paper begins to remedy that lack in the simplest case by presenting a peridynamic equivalent to an Euler-Bernoulli beam.
In addition to directly modeling a thin beam in bending, the simple beam case lays the theoretical framework for more complex peridynamic beam, plate, and shell bending models.
Because many analyses of interest are partly or wholly comprised of these types of features, their development is an important addition to the capabilities of peridynamic analysis.
%
%The second section of this paper provides a brief introduction to peridynamics, including state based models.
%The third section presents the state based beam model and demonstrates equivalence to classical Euler-Bernoulli beam theory in the limit of shrinking nonlocality.
%The fourth section demonstrates the model's equivalence to Eringen's nonlocal elasticity for small peridynamic horizons, and compares other gradient models.
%The fifth section demonstrates the beam model with simple test cases.
The remainder of this introduction reviews other nonlocal work and provides a brief introduction to peridynamics, including state based models.
\Cref{sec:ModelDev} presents the state based beam model and demonstrates equivalence to classical Euler-Bernoulli beam theory in the limit of shrinking nonlocality.
\Cref{sec:EringenCompare} demonstrates the model's equivalence to Eringen's nonlocal elasticity for small peridynamic horizons, and compares other gradient models.
\Cref{sec:Numerical} demonstrates the beam model with simple test cases.
%
\subsection{Nonlocal Beam Models}
\label{sec:NLbeams}
%
Nonlocal elasticity generally allows for forces at a point that are dependent on the material configuration of an entire body, rather than the configuration at that point.\cite{eringen1972nonlocal}
While long-range forces are obvious at the molecular model, material at larger scales is conventionally modeled as though internal forces are local or contact forces.\cite{kroner1967elasticity}
The result of such approximation is accurate for deformations that are homogenous, but introduces some inaccuracy for inhomogenous deformations like the propagation of waves with short wavelengths.
One way to distinguish between homogenous and inhomogenous deformations is to incorporate higher-order gradients of deformation.
While stress in classical elasticity is a function of the (first) gradient of deformation, Eringen's formulation of a nonlocal modulus in \cite{eringen1983differential} approximates a weighted sum of the first and second order gradients.
This introduces a length scale to the model and has the effect of smearing out local deformation inhomogeneities over the surrounding material, while maintaining the conventional result for homogenous deformations.

Previous work in the nonlocal mechanics of beams is motivated by the observed stiffening of nanoscale cantilevers.
Challamel and Wang demonstrate in \cite{Challamel2008small} that Eringen nonlocal elasticity cannot reproduce the scale stiffening, but that stiffening does result from other gradient-elastic models and models incorporating nonlocal curvature.
Because all of these models incorporate higher-order gradients of deformation, they impose stronger continuity requirements than classical elasticity, and are unsuitable for discontinuous displacements.
Because the gradients are evaluated locally, gradient models are called \textit{weakly nonlocal}.

%
\subsection{Peridynamics}
\label{sec:PDintro}
The term \textit{peridynamic} was coined by Silling in \cite{silling2000reformulation} and alludes to the fact that the force at a point is affected by nearby material configuration.
In contrast to gradient models, the peridynamic model is \textit{strongly nonlocal} and casts material behavior at a point as the \textit{integral} equation 
%
\begin{equation}
\label{eq:PDCoPV}
\rho(\mathbf{x})\ddot{\mathbf{u}}(\mathbf{x}) = \int_\Omega \mathbi{f}(\mathbf{x},\mathbf{q}) dV_\mathbf{q}  + \mathbf{b}(\mathbf{x}) \notag
\end{equation}
%
rather than the classical \textit{differential} equation.
Instead of the divergence of stress, we have the integral of a ``force" function $\mathbi{f}$ of the position vector $\mathbf{x}$ and the position vector $\mathbf{q}$ of a point within the body domain $\Omega$. 
This force function may depend on \(\mathbf{x}\), \(\mathbf{q}\), their deformed positions, the original and deformed positions of other points in \(\Omega\), history, etc.

@@

\label{sec:ModelDev}

@@

\label{sec:EnergyEq}

@@

\subsection{Weighting Function and Perfect Plasticity}
\label{sec:WeightFunction}
The weighting function \(\omega(\xi)\) describes the relative contribution of each bond-pair, and can be defined according to physical or mathematical considerations. 
Consider an Euler beam in bending with curvature \(\kappa\). 
Fibers running parallel to the neutral axis of the beam are stretched in proportion to their distance from the neutral axis, with strain \(\epsilon = y\kappa\). 
If the fibers are linearly elastic, then the axial stress at each location is \(\sigma = E\epsilon = Ey\kappa\), and the contribution to supported moment \(dM = \kappa E y^2 dA\). 
By comparing the formulations for the moments carried by the Euler beam in \cref{fig:EulerBending} and those of the bond-pair beam in \cref{fig:BPBending}, we see some definite parallels.
%
\begin{figure}[h]
\subinputfrom{\diagrampath}{EulerBending.eps_tex}
\caption{Euler beam moment contribution}
\label{fig:EulerBending}
\end{figure}
%%
\begin{figure}[h]
\subinputfrom{\diagrampath}{BondPairBending.eps_tex}
\caption{Bond-pair moment contribution}
\label{fig:BPBending}
\end{figure}
%
\begin{align}
M_\text{classical}&=\int_{-\frac{t}{2}}^{\frac{t}{2}} \sigma \; y \; dA && = \int_{-\frac{t}{2}}^{\frac{t}{2}} E \kappa \; y^2 \; b(y) dy\notag \\
%
M_\text{PD}&=\int_{-\delta}^{\delta} \vstate{T}{}{\xi}\; \xi \; d\xi & = \int_{-\delta}^{\delta} \alpha \frac{\sin(\Delta\theta)}{|\xi|} \; \xi \; \omega(\xi) d\xi\: & \approx \int_{-\delta}^{\delta} \alpha \kappa |\xi| \; \omega(\xi) d\xi\notag
\end{align}
%
The term \(y\) is the distance from the beam's neutral axis and \(b(y)\) is the width of the beam at that distance from the neutral axis. 
The similarity between classical and peridynamic moment formulations suggests a possible formulation for the weighting function:
%
\begin{equation}
\omega(\xi) = |\xi| b\left(y\right) \text{at}\; y=\frac{\xi}{\delta} \frac{t}{2}\notag
\end{equation}
%
\begin{figure}[h]
\subinputfrom{\diagrampath}{WeightProfile.eps_tex}
\caption{A cross-section dependent weight function}
\label{fig:WeightProfile}
\end{figure}
%
This weight function analogizes the relative contributions of bond pairs of different lengths to the relative contributions of fibers at different distances from the centerline. 
An example is shown in \cref{fig:WeightProfile}.
While this weighting function offers no advantages over a uniform weight function in the case of the linearly elastic beam, it offers a way to model advancing plasticity.

@@

\label{sec:EringenCompare}

@@

\section{Numerical Simulation}
\label{sec:Numerical}
\subsection{Discretized Model}
\label{sec:Discretized}
For analysis purposes, it is often useful to represent a continuum part with a discrete version. 
Discretizing the bond-pair model is primarily matter of exchanging integrals for sums. 
%
\begin{align}
%\label{eq:discreteBeamw}
w = \alpha \left[1+\cos(\theta(\vstate{Y}{}{\xi},\vstate{Y}{}{\mathbf{-\xi}})) \right] \approx \frac{\alpha}{2}\left(\frac{y(x+\xi)-2y(x)+y(x-\xi)}{\xi}\right)^2 \notag
\end{align}
%
\begin{align}
\label{eq:discretebeam}
\alpha &= \frac{c\; \Delta x}{m} ;\; c= EI ;\; m=\sum_{i=1}^n \omega(\xi_i)\xi_i^2 \implies \nonumber \\
W&=\Delta x \sum_{i=1}^n \frac{EI}{2}\left(\frac{y(x+\xi_i)-2y(x)+y(x-\xi_i)}{\xi_i}\right)^2\notag
\end{align}
%

Discretizing the bond-pair model as proposed above requires that nodes be evenly spaced along the entire beam, otherwise the displacement \(y(x-\xi_i)\) is ill-defined. 
For this reason, the discretization does not allow for areas of higher and lower ``resolution''. 
\subsection{Results}
\label{sec:Results}
The simplest test case for this model is a linear-elastic beam with a square profile.
The models below were constrained to purely vertical movement of each point in order to reduce the total number of degrees of freedom, speeding evaluation and improving convergence.
For comparison, equivalent models are created and analyzed in Abaqus 6.12 to verify simple cases.
Even a coarse discretization successfully reproduces the elastically deformed beam shape in \cref{fig:eppt_g100}.

@@

\section{Conclusion}
\label{sec:Conclusion}
